The Astropy Project: \\ Sustaining and Growing a Community-oriented Open-source Project and the Latest Major Release (v5.0) of the Core Package

### 2022-03-11 src/ms.tex
new figure and words

\subsection{Contributor base} \label{sec:project-contributors}

@@

Overview and statistics of contributors. Highlight changes since v2.0.

\begin{figure}
    \begin{centering}
      \includegraphics[width=0.5\textwidth]{commits_committers_per_month.pdf}
        \caption{
            todo
        }
        \label{fig:contributor-summary:per-month}
    \end{centering}
\end{figure}

\begin{figure}
    \begin{centering}
      \includegraphics[width=\textwidth]{Ncommitters_vs_frac_commits.pdf}
        \caption{
            todo
        }
        \label{fig:contributor-summary:Ncommitters}
    \end{centering}
\end{figure}

\begin{figure}
    \begin{centering}
      \includegraphics[width=0.75\textwidth]{top_contributors_career.pdf}
        \caption{
            todo
        }
        \label{fig:contributor-summary:top}
    \end{centering}
\end{figure}

>>>>

\subsection{Core package contributor base} \label{sec:project-contributors}

@@

% TODO: move this subsection to "updates to core package" section above?

\begin{figure}[ht]
    \begin{centering}
      \includegraphics[width=0.8\textwidth]{Ncommitters_vs_frac_commits.pdf}
        \caption{
            todo
        }
        \label{fig:contributor-summary:Ncommitters}
    \end{centering}
\end{figure}

As described in the introduction above, the \astropy Project is broader than the
core \astropypkg package as it also represents a number of core infrastructure
packages, affiliated packages, educational initiatives, and user communities.
In total, well over 1500 people have contributed in some way to the Project,
whether, for example, through contributing code or documentation, opening issues
on \github, providing infrastructure support, writing tutorials, or helping to
coordinate workshops that relate to Astropy.
The vast majority of these contributors and participants have interacted as
volunteers choosing to engage with the open-source and open-development software
community.
However, while the total number of contributors is large, the number of
contributions and bulk of the maintenance of the \astropy Project components has
primarily been sustained by a small number of heavily-engaged or
explicitly-employed contributors \citep[see, e.g., ][]{astropy:2018}, as is the
case in many large, open-source software communities \citep[e.g.,][]{numpy}.

As an example of this effect within the \astropy Project, here we focus on
contributions made to the \astropypkg core package.
Figure~\ref{fig:contributor-summary:Ncommitters} shows the number of
contributors who have contributed a given fraction of the total number of Git
commits, separated into three separate ``eras'' in the history of the
Project.\footnote{Note, however, that measuring these statistics in terms of Git
commits is not a perfect way of encapsulating ``effort,'' but we use this metric
for its simplicity and clear definition.}
The annotations show that, within the early (July 2011--January 2015), mid
(January 2015--June 2018), and recent (June 2018--November 2021) eras, 90\% of
the commits to the \astropypkg core package have come from 18, 34, and 44
contributors.
While the trend is encouraging --- a larger pool of contributors are
contributing more, fractionally --- this emphasizes that, still, the vast
majority of the work in \astropypkg is done my a small subset of the
contributors.

\begin{figure}[ht]
    \begin{centering}
      \includegraphics[width=0.5\textwidth]{commits_committers_per_month.pdf}
        \caption{
            todo
        }
        \label{fig:contributor-summary:per-month}
    \end{centering}
\end{figure}

\begin{figure}[ht]
    \begin{centering}
      \includegraphics[width=0.75\textwidth]{top_contributors_career.pdf}
        \caption{
            todo
        }
        \label{fig:contributor-summary:top}
    \end{centering}
\end{figure}
